### 2024-10-25 Paper/paper.tex

    \import{Tables}{leader_sumstats.tex}

@@

    \begin{figure}[ht!]
    \centering
        \caption{Outcomes Over Time}
        \includegraphics[width=\textwidth]{Objects/weighted_read_mort_adjusted_graph.pdf}
        \label{fig:weighted_read_mort_graph}
    \end{figure}

@@

     \begin{figure}
     \caption{Readmission Rate Synthetic Difference in Differences Results}
     \centering
          \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and All NFP}
         \includegraphics[width=\textwidth]{Objects/read_fp_nfp_synth_graph.pdf}
         \label{fig:read_synth_plota}
     \end{subfigure}%
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and NFP w/ MD Exec}
         \includegraphics[width=\textwidth]{Objects/read_fp_md_synth_graph.pdf}
         \label{fig:read_synth_plotb}
     \end{subfigure}%
     \vspace{5mm}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and NFP w/out MD Exec}
         \includegraphics[width=\textwidth]{Objects/read_fp_nomd_synth_graph.pdf}
         \label{fig:read_synth_plotc}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{NFP w/ and w/out MD Exec}
         \includegraphics[width=\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
         \label{fig:read_synth_plotd}
     \end{subfigure}
        \label{fig:read_synth_plot}
    \end{figure}

@@

     \begin{figure}
     \caption{Mortality Rate Synthetic Difference in Differences Results}
     \centering
          \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and All NFP}
         \includegraphics[width=\textwidth]{Objects/mort_fp_nfp_synth_graph.pdf}
         \label{fig:mort_synth_plota}
     \end{subfigure}%
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and NFP w/ MD Exec}
         \includegraphics[width=\textwidth]{Objects/mort_fp_md_synth_graph.pdf}
         \label{fig:mort_synth_plotb}
     \end{subfigure}%
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{For-Profit and NFP w/out MD Exec}
         \includegraphics[width=\textwidth]{Objects/mort_fp_nomd_synth_graph.pdf}
         \label{fig:mort_synth_plotc}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{NFP w/ and w/out MD Exec}
         \includegraphics[width=\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
         \label{fig:mort_synth_plotd}
     \end{subfigure}
        \label{fig:mort_synth_plot}
    \end{figure}

@@

     \import{Tables}{change_analysis.tex}

@@

\begin{figure}[ht!]
    \centering
    \caption{Executive Team Changes Over Time}
    \includegraphics[width=\textwidth]{Objects/change_graph.pdf}
    \label{fig:change_graph}
\end{figure}

@@

\import{Tables}{forprofit_readmort_fullsample.tex}

@@

\import{Tables}{MD_noMD_readmort_fullsample.tex}

@@

\begin{figure}[ht!]
     \caption{TWFE Event Study Results}
     \centering
          \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{}
         \includegraphics[width=\textwidth]{Objects/read_forprofit_es_graph.pdf}
         \label{fig:es_plota}
     \end{subfigure}%
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{}
         \includegraphics[width=\textwidth]{Objects/read_MD_es_graph.pdf}
         \label{fig:es_plotb}
     \end{subfigure}%
     \vspace{5mm}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{}
         \includegraphics[width=\textwidth]{Objects/mort_forprofit_es_graph.pdf}
         \label{fig:es_plotc}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{}
         \includegraphics[width=\textwidth]{Objects/mort_MD_es_graph.pdf}
         \label{fig:es_plotd}
     \end{subfigure}
        \label{fig:es_plot}
    \end{figure}

@@

\import{Tables}{forprofit_readmort_synth.tex}

\import{Tables}{MD_noMD_readmort_synth.tex}

>>>>

    %\import{Tables}{leader_sumstats.tex}

@@

    %\begin{figure}[ht!]
    %\centering
        %\caption{Outcomes Over Time}
        %\includegraphics[width=\textwidth]{Objects/weighted_read_mort_adjusted_graph.pdf}
        %\label{fig:weighted_read_mort_graph}
    %\end{figure}

@@

    %  \begin{figure}
    %  \caption{Readmission Rate Synthetic Difference in Differences Results}
    %  \centering
    %       \begin{subfigure}[b]{0.45\textwidth}
    %      \centering
    %      \caption{For-Profit and All NFP}
    %      \includegraphics[width=\textwidth]{Objects/read_fp_nfp_synth_graph.pdf}
    %      \label{fig:read_synth_plota}
    %  \end{subfigure}%
    %  \hfill
    %  \begin{subfigure}[b]{0.45\textwidth}
    %      \centering
    %      \caption{For-Profit and NFP w/ MD Exec}
    %      \includegraphics[width=\textwidth]{Objects/read_fp_md_synth_graph.pdf}
    %      \label{fig:read_synth_plotb}
    %  \end{subfigure}%
    %  \vspace{5mm}
    %  \hfill
    %  \begin{subfigure}[b]{0.45\textwidth}
    %      \centering
    %      \caption{For-Profit and NFP w/out MD Exec}
    %      \includegraphics[width=\textwidth]{Objects/read_fp_nomd_synth_graph.pdf}
    %      \label{fig:read_synth_plotc}
    %  \end{subfigure}
    %  \hfill
    %  \begin{subfigure}[b]{0.45\textwidth}
    %      \centering
    %      \caption{NFP w/ and w/out MD Exec}
    %      \includegraphics[width=\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
    %      \label{fig:read_synth_plotd}
    %  \end{subfigure}
    %     \label{fig:read_synth_plot}
    % \end{figure}

@@

    %  \begin{figure}
    %  \caption{Mortality Rate Synthetic Difference in Differences Results}
    %  \centering
    %       \begin{subfigure}[b]{0.45\textwidth}
    %      \centering
    %      \caption{For-Profit and All NFP}
    %      \includegraphics[width=\textwidth]{Objects/mort_fp_nfp_synth_graph.pdf}
    %      \label{fig:mort_synth_plota}
    %  \end{subfigure}%
    %  \hfill
    %  \begin{subfigure}[b]{0.45\textwidth}
    %      \centering
    %      \caption{For-Profit and NFP w/ MD Exec}
    %      \includegraphics[width=\textwidth]{Objects/mort_fp_md_synth_graph.pdf}
    %      \label{fig:mort_synth_plotb}
    %  \end{subfigure}%
    %  \hfill
    %  \begin{subfigure}[b]{0.45\textwidth}
    %      \centering
    %      \caption{For-Profit and NFP w/out MD Exec}
    %      \includegraphics[width=\textwidth]{Objects/mort_fp_nomd_synth_graph.pdf}
    %      \label{fig:mort_synth_plotc}
    %  \end{subfigure}
    %  \hfill
    %  \begin{subfigure}[b]{0.45\textwidth}
    %      \centering
    %      \caption{NFP w/ and w/out MD Exec}
    %      \includegraphics[width=\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
    %      \label{fig:mort_synth_plotd}
    %  \end{subfigure}
    %     \label{fig:mort_synth_plot}
    % \end{figure}

@@

     % \import{Tables}{change_analysis.tex}

@@

% \begin{figure}[ht!]
%     \centering
%     \caption{Executive Team Changes Over Time}
%     \includegraphics[width=\textwidth]{Objects/change_graph.pdf}
%     \label{fig:change_graph}
% \end{figure}

@@

%\import{Tables}{forprofit_readmort_fullsample.tex}

@@

%\import{Tables}{MD_noMD_readmort_fullsample.tex}

@@

% \begin{figure}[ht!]
%      \caption{TWFE Event Study Results}
%      \centering
%           \begin{subfigure}[b]{0.45\textwidth}
%          \centering
%          \caption{}
%          \includegraphics[width=\textwidth]{Objects/read_forprofit_es_graph.pdf}
%          \label{fig:es_plota}
%      \end{subfigure}%
%      \hfill
%      \begin{subfigure}[b]{0.45\textwidth}
%          \centering
%          \caption{}
%          \includegraphics[width=\textwidth]{Objects/read_MD_es_graph.pdf}
%          \label{fig:es_plotb}
%      \end{subfigure}%
%      \vspace{5mm}
%      \hfill
%      \begin{subfigure}[b]{0.45\textwidth}
%          \centering
%          \caption{}
%          \includegraphics[width=\textwidth]{Objects/mort_forprofit_es_graph.pdf}
%          \label{fig:es_plotc}
%      \end{subfigure}
%      \hfill
%      \begin{subfigure}[b]{0.45\textwidth}
%          \centering
%          \caption{}
%          \includegraphics[width=\textwidth]{Objects/mort_MD_es_graph.pdf}
%          \label{fig:es_plotd}
%      \end{subfigure}
%         \label{fig:es_plot}
%     \end{figure}

@@

%\import{Tables}{forprofit_readmort_synth.tex}

%\import{Tables}{MD_noMD_readmort_synth.tex}

### 2024-10-25 Paper/paperdraft_Oct2024.tex

    In corporate environments, executives are considered crucial drivers of a company's practices and performance, as they are heavily involved in the culture, visions, and strategies of the firm. Executive heterogeneity has been linked to various actions and outcomes of firms; characteristics such as gender, age, and background are correlated with financial performance and behaviors such as employee diversity, wages, aggressiveness in mergers and acquisitions, and other corporate strategies.\footnote{\cite{bertrand2003managing}; \cite{matsa2011chipping}; \cite{matsa2011chipping}; \cite{adams2009women}; \cite{ahern2012changing}; \cite{flabbi2019female}; \cite{benmelech2015military}; \cite{custodio2013generalists}; \cite{frydman2019rising}} However, there are still settings and relevant characteristics for which the relationship between management and firm behavior is not well understood.

    In this paper, I study whether type of executive training affects firm behavior in the context of US hospitals. In the US, health care makes up almost 20\% of spending (\cite{CMS_National_Health_Expenditure}), and a major goal of policymakers is to improve the quality of health care without increasing costs. Yet, little is known about how hospital managers affect health care value, and it is unclear whether existing findings on executive characteristics can be extrapolated to the health care, and particularly hospital, context. Existing literature is generally limited to large, publicly traded firms, whereas a majority of US hospitals are nonprofit with presumably different objectives. In particular, health care executives have a particularly unique trade-off between profit and patient welfare, which varies from the types of firms studied in existing literature.

    A characteristic relevant to this context is whether or not hospital executives have clinical training. Physicians can bring a unique combination of clinical and administration expertise, and they have insights that can potentially benefit their employees and lead to better quality (\cite{Stajduhar_2023}, \cite{Ahmed_2022}). However, hiring physician executives is controversial, as physicians are not always trained in business management, which can lead to poor practices and lower quality (\cite{HarvardBusinessReview2018}). Thus, I estimate how having clinically trained executives in hospital leadership affects hospital outcomes. I scrape executive names and titles of nonprofit hospitals in the US from publicly available Tax Form 990s. Combining this with other public hospital data, I construct a novel data set of nonprofit hospitals from 2010-2014 containing information on both executive team characteristics and quality metrics. While limited to nonprofits due to data availability, this sample of hospitals is an important driver of health care in the US. Private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds. For comparison, for-profits make up 36\% of hospitals and staff 107 beds on average (\cite{ASPE_2023}).

@@

@@

    One hypothesis that could explain the differential response due to clinical leadership is how profit-driven a clinician is compared to other training backgrounds. To demonstrate this, consider hospital behavior under the enactment of pay-for-performance policies, where hospitals face a trade-off between profit and societal/patient benefit. Pay-for-performance incentives essentially incorporate quality directly into a hospital's profit function (\cite{dranove2011health}), whereas a fee-for-service system only incentivizes quantity. A hospital that places all value on profit will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.

    To formalize this intuition, I specify the hospital objective function as a weighted average of profit and societal (or patient) outcomes and derive characteristics of the optimal quality. For simplicity, I abstract away from quality having a direct effect of quantity and price. Without any financial incentives to increase quality, hospital quality is decided by:

    \noindent where $\pi(\theta)$ is a profit function depending on $\theta$, and the relationship between profit and $\theta$ depends on pay-for-performance incentives. Additionally, $u(.)$ is extra utility gained from quality, which is increasing and concave in $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs. societal benefit. There is an implicit stay-open condition throughout the model. Taking the first order condition yields

    \noindent where marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$:

    $$\frac{du'(\theta)}{d\alpha} = \frac{-1}{(1-\alpha)^2}\pi'(\theta)$$

    First, assume the hospital is in a strictly fee-for-service setting, where quality benefits patients, but does not directly provide profit to the hospital through reimbursements. Then, assuming there is a positive cost to increasing quality, $\pi(\theta)$ is decreasing in $\theta$. Then $\frac{du'(\theta)}{d\alpha} > 0$, revealing that in a fee-for-service setting, the more weight placed on profit, the lower quality of care chosen by the hospital.

    Second, assume the hospital is in a strictly pay-for-performance setting, where quality benefits patients and directly increases profit. Then, assuming the marginal financial benefit of increasing quality is greater than the marginal financial cost, $\pi(\theta)$ is increasing in $\theta$. Then $\frac{du'(\theta)}{d\alpha} < 0$, revealing that in a pay-for-performance setting, the more weight placed on profit, the higher quality of care chosen by the hospital. I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_{p4p}$ is quality under pay-for-performance incentives and $\theta_{ffs}$ is quality under fee-for-service only. That is,

    Hence, this model predicts that change in quality depends on how much weight the hospital places on profit vs. societal benefit. Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit. In reality, $\alpha$ is unknown, except in the case of for-profits, who have revealed that they have a high $\alpha$ by their ownership status. Clinically trained executives could also affect this $\alpha$, as doctors commonly take a hippocratic oath to put patients first.

    \subsection{Estimation}

>>>>

    In corporate environments, executives are considered crucial drivers of a company's practices and performance, as they are heavily involved in the culture, visions, and strategies of the firm. Executive heterogeneity has been linked to various actions and outcomes of firms; characteristics such as gender, age, and background are correlated with financial performance and behaviors such as employee diversity, wages, aggressiveness in mergers and acquisitions, and other corporate strategies.\footnote{\cite{bertrand2003managing}; \cite{matsa2011chipping}; \cite{matsa2011chipping}; \cite{adams2009women}; \cite{ahern2012changing}; \cite{flabbi2019female}; \cite{benmelech2015military}; \cite{custodio2013generalists}; \cite{frydman2019rising}} However, one aspect of firm leadership that is not well understood is how different leaders may interact with changing policy incentives differently.

    In this paper, I study whether type of executive training affects firm behavior in the context of US hospitals. In the US, health care makes up almost 20\% of spending (\cite{CMSNationalHealthExpenditure}), and a major goal of policymakers is to improve the quality of health care without increasing costs. Yet, little is known about how hospital managers affect health care value, and it is unclear whether existing findings on executive characteristics can be extrapolated to the health care, and particularly hospital, context. Existing literature is generally limited to large, publicly traded firms, whereas a majority of US hospitals are nonprofit with presumably different objectives and policy incentives.

    A characteristic relevant to this context is whether or not hospital executives have clinical training. Physicians can bring a unique combination of clinical and administration expertise, and they have insights that can potentially benefit their employees and lead to better quality (\cite{Stajduhar_2023}, \cite{Ahmed_2022}). However, hiring physician executives is controversial, as physicians are not always trained in business management, which can lead to poor practices and lower quality (\cite{HarvardBusinessReview2018}). Thus, I estimate how clinically trained executives in hospital leadership affects hospital outcomes. I scrape executive names and titles of nonprofit hospitals in the US from publicly available Tax Form 990s. Combining this with other public hospital data, I construct a novel data set of nonprofit hospitals from 2010-2014 containing information on both executive team characteristics and quality metrics. While limited to nonprofits due to data availability, this sample of hospitals is an important driver of health care in the US. Private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds. For comparison, for-profits make up 36\% of hospitals and staff 107 beds on average (\cite{ASPE_2023}).

@@

    \section{Theoretical Framework}

    Consider a simplified model of hospital objectives, in which hospitals choose a level of quality of care that maximizes an objective function capturing the trade-off between profit and patient welfare. That is,

    \noindent where $\pi(\theta)$ is a profit function depending on quality $\theta$, $u(.)$ is extra utility gained from quality, which is increasing and concave in $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs. societal benefit. There is an implicit stay-open condition throughout the model.

    In this paper, I consider hospital choice of quality under pay-for-performance incentives, which essentially incorporate quality directly into a hospital's profit function (\cite{dranove2011health}). Alternatively, a fee-for-service system only incentivizes quantity. A hospital that places all value on profit will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives. I will now formalize this intuition mathematically. For simplicity, I abstract away from quality having a direct impact on quantity of patients or price.\footnote{As long as the relationship between quality and quantity/prove are not dependent on clinical experience on the leadership team, this does not change the conclusion of the model.} Taking the first order condition yields:

    \noindent where marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$,

    $$\frac{du'(\theta)}{d\alpha} = \frac{-1}{(1-\alpha)^2}\pi'(\theta).$$

    Now, consider incentives when the hospital is in a strictly fee-for-service environment. That is, quality benefits patients, but hospitals do not directly derive profit from having a high quality. Then, assuming there is a positive cost to increasing quality, $\pi(\theta)$ is decreasing in $\theta$. Then $\frac{du'(\theta)}{d\alpha} > 0$. Using the Implicit Function, I can write $\frac{d\theta}{d\alpha}$ as 
    $\frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta}<0$, revealing that in a fee-for-service setting, the more weight the hospital places on profit, the lower quality of care chosen.

    Next, assume the hospital is in a strictly pay-for-performance setting, where quality benefits patients and directly increases profit. Then, assuming the marginal financial benefit of increasing quality is greater than the marginal financial cost, $\pi(\theta)$ is increasing in $\theta$. Then $\frac{du'(\theta)}{d\alpha} < 0$, revealing that in a pay-for-performance setting, the more weight placed on profit, the higher quality of care chosen by the hospital. I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_{p4p}$ is quality under pay-for-performance incentives and $\theta_{ffs}$ is quality under fee-for-service only. That is,

    Hence, this model predicts that change in quality depends on how much weight the hospital places on profit vs. societal benefit. Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit. Therefore, for leadership characteristics that are plausibly correlated with how much weight is placed on patient welfare versus quality, such as clinical training, we expect to see a differential response to pay-for-performance incentives being enacted.

@@
